Verification in Isabelle/HOL of a Set-based Algorithm for Computing Strongly Connected Components

### 2022-09-20 cpp/submission.tex
Summary of proof effort in a table

\newcommand{\prog}[1]{\textit{#1}}
\renewcommand{\SS}{\mathcal{S}}

@@

In the statement of the lemma, \prog{wf} represents the well-foundedness predicate, which is predefined in the Isabelle standard library. The lemma is proved by embedding the argument tuples into triples $(a,b,c)$ where $a$ is the set of unvisited vertices, $b$ is the set of edges between vertices that have not been followed, and $c$ is $0$ if the triple represents an argument of \prog{dfs} and $1$ if it is an argument of \prog{dfss}. These triples can be ordered by the strict subset order for the two first component and the predecessor relation on natural numbers for the third one, and Isabelle proves automatically that the product of these orders is well-founded. Finally, \prog{dfs-dfss-term} can be embedded as an inverse image in this well-founded relation, and is therefore itself well-founded.

@@

\textbf{Summarize the overall effort of the proof in number of lines (definitions / proofs) for each main category.}

>>>>

\newcommand{\prog}[1]{\textit{#1}}
\renewcommand{\SS}{\mathcal{S}}
\usepackage{array}
\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}

@@

In the statement of the lemma, \prog{wf} represents the well-foundedness predicate, which is predefined in the Isabelle standard library. The lemma is proved by embedding the argument tuples into triples $(a,b,c)$ where $a$ is the set of unvisited vertices, $b$ is the set of edges between vertices that have not been followed, and $c$ is $0$ if the triple represents an argument of \prog{dfs} and $1$ if it is an argument of \prog{dfss}. These triples can be ordered by the strict subset order for the two first components and the predecessor relation on natural numbers for the third one, and Isabelle proves automatically that the product of these orders is well-founded. Finally, \prog{dfs-dfss-term} can be embedded as an inverse image in this well-founded relation, and is therefore itself well-founded.

@@

We summarize the overall effort of the proof in number of lines of the Isabelle theory (definitions / proofs) in a table for each main category.

\begin{center}
  \small
  \begin{tabular}{|M{.125\textwidth}|M{.075\textwidth}|m{.21\textwidth}|}
    \hline
    Category & Lines & Description\\
    \hline
    Lists ordering  & 102-240 & We use a relation of precedence in lists that determines a partial order on the elements that appear in a list, provided those elements have unique occurrences. In particular, stacks are represented as lists.\\
    \hline
    Graphs & 251-424 & We prove a few lemmas on basic concepts in graphs. In particular, we define reachability and SCCs and prove some auxiliary lemmas on them.\\
    \hline
    SCC algorithm & 444-492 & We formalize the algorithm by defining functions \prog{unite}, \prog{dfs} and \prog{dfss}.\\
    \hline
    Invariants & 500-886 & We formalize the main invariants, i.e. \prog{wf-env} and the pre- and post-conditions for both functions \prog{dfs} and \prog{dfss} and prove some direct consequences of these invariants.\\
    \hline
    Partial correctness & 897-3087 & We prove partial correctness by proving that the pre-conditions imply the post-conditions.\\
    \hline
    Termination & 3140-3413 & We prove termination using the fact that at each call, either a new vertex is visited or a new successor is added to the set of visited successors of some vertex.\\
    \hline
    Total correctness & 3419-3423 & We prove total correctness using termination and partial correctness.\\
    \hline
  \end{tabular}
\end{center}
